---
title: "RNA-seq miniCURE"
date: "`r format(Sys.time(), '%B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Description about Course/Book."
favicon: assets/dasl_favicon.ico
output:
    bookdown::word_document2:
      toc: true
---

# About this Course {-}

This miniCURE allows students to develop a project using RNA-seq datasets to introduce how model organisms, high-throughput sequencing, and the scientific process are used in both basic and clinical research.

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1rWH7VTcPV1juH0E9NI-X6evMIKzgn1MQKlf_CRzT73w/edit#slide=id.g1f734f625ef_0_30")
```

### Skills Level {- .unlisted}

::: {.notice}
_Genetics_  
**Novice**: Introduction to [central dogma of molecular biology](https://openstax.org/books/biology-2e/pages/15-1-the-genetic-code)

_Programming skills_  
**Novice**: No programming experience needed
:::

### Format {- .unlisted}

- **Class Type:** Lab (computer based)
- **Lesson Length:** Multiple class periods (4-5 lab periods)
    - Scientific Literature
    - Model Organisms and Databases
    - RNA-seq Analysis
    - Differential Gene Expression
    - *Kickstart Project Work

*Kickstart Project Work is flexible - it is enormously helpful for students to to have some guided lab time to work on the project, but this can be overlapped with a wet-lab protocol with a lot of downtime, such as PCR.
    
    
### Learning Goals {-}

- Learn about model organisms
- Explore high-throughput sequencing datasets
- Practice the [scientific process](https://undsci.berkeley.edu/understanding-science-101/how-science-works/)

### C-MOOR Collection {- .unlisted}

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "about_c-moor.Rmd",
  repo_name = "C-MOOR/C-MOOR_Template"
)
```
