---
title: "RNA-seq miniCURE"
date: "`r format(Sys.time(), '%B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Description about Course/Book."
favicon: assets/dasl_favicon.ico
output:
    bookdown::word_document2:
      toc: true
---

# About this Course {-}

This miniCURE allows students to develop a project using RNA-seq datasets to introduce how model organisms, high-throughput sequencing, and the scientific process are used in both basic and clinical research.

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1rWH7VTcPV1juH0E9NI-X6evMIKzgn1MQKlf_CRzT73w/edit#slide=id.g1f734f625ef_0_30")
```

### Skills Level {- .unlisted}

::: {.notice}
_Genetics_  
**Novice**: Introduction to [central dogma of molecular biology](https://openstax.org/books/biology-2e/pages/15-1-the-genetic-code)

_Programming skills_  
**Novice**: No programming experience needed
:::

### Learning Goals {-}

- **Navigate biological databases** to access the wealth of available scientific knowledge and data.
- **Analyze data (i.e. create and interpret plots)** to generate novel conclusions.
- **Cultivate an “-omics” perspective**, integrating multiple sources of information and wielding computational tools to enable a holistic approach
- **Practice the scientific process**, identifying avenues for research, designing experiments, analyzing data, and integrating results into the broader scientific discourse.

### Core Competencies {- .unlisted}

This activity addresses the following core concepts and competencies:

```{r, echo=FALSE}
competencies <- tibble::tribble(
  ~source, ~competencies,
  
    "**Vision and Change**",  paste(
                              "Concepts\n",
                              "<ul>",
                              # "<li>**1**: EVOLUTION: The diversity of life evolved over time by processes of mutation, selection, and genetic change.</li>",
                              # "<li>**2**: STRUCTURE AND FUNCTION: Basic units of structure define the function of all living things.</li>",
                              "<li>**3**: INFORMATION FLOW, EXCHANGE, AND STORAGE</li>",
                              # "<li>**4**: PATHWAYS AND TRANSFORMATIONS OF ENERGY AND MATTER: Biological systems grow and change by processes based upon chemical transformation pathways and are governed by the laws of thermodynamics.</li>",
                              # "<li>**5**: SYSTEMS: Living systems are interconnected and interacting.</li>",
                              "</ul>",
                              
                              "Competencies\n",
                              "<ul>",
                              "<li>**1**: Ability to apply the process of science</li>",
                              "<li>**2**: Ability to use quantitative reasoning</li>",
                              # "<li>**3**: Ability to use modeling and simulation</li>",
                              # "<li>**4**: Ability to tap into the interdisciplinary nature of science</li>",
                              "<li>**5**: Ability to communicate and collaborate with other disciplines</li>",
                              # "<li>**6**: Ability to understand the relationship between science and society</li>",
                              "</ul>",
                              sep="\n"),
  
    "**Genetics**",           paste(
                              "Concepts\n",
                              "<ul>",
                              # "<li>How is DNA organized?</li>",
                              # "<li>What are the molecular components and mechanisms necessary to preserve and duplicate an organism’s genome?</li>",
                              # "<li>What are the mechanisms by which an organism’s genome is passed on to the next generation?</li>",
                              # "<li>How can one deduce information about genes, alleles, and gene functions from analysis of genetic crosses and patterns of inheritance?</li>",
                              # "<li>How does the phenomenon of linkage affect the assortment of alleles during meiosis?</li>",
                              "<li>How is genetic information expressed so it affects an organism’s structure and function?</li>",
                              # "<li>How can gene activity be altered in the absence of DNA changes?</li>",
                              # "<li>How do genes and genomes control changes in an organism’s structure and function throughout its life cycle?</li>",
                              # "<li>How do different types of mutations affect genes and the corresponding mRNAs and proteins?</li>",
                              # "<li>What are the processes that can affect the frequency of genotypes and phenotypes in a population over time?</li>",
                              "<li>How do the results of molecular genetic studies in model organisms help us understand aspects of human genetics and genetic diseases?</li>",
                              "<li>What experimental methods are commonly used to analyze gene structure, gene expression, gene function, and genetic variants?</li>",
                              "</ul>",
                              
                              "Competencies\n",
                              "<ul>",
                              "<li>Locate, read, and comprehend primary literature research papers on genetics topics</li>",
                              "<li>Implement observational strategies to formulate a question</li>",
                              "<li>Generate testable hypotheses</li>",
                              # "<li>Design an experiment using appropriate controls and appropriate sample sizes</li>",
                              # "<li>Gather and evaluate experimental evidence, including qualitative and quantitative data</li>",
                              # "<li>Apply statistical methods when analyzing their data, and use patterns to construct a model</li>",
                              "<li>Generate and interpret graphs displaying experimental results</li>",
                              "<li>Critique large data sets and use bioinformatics to assess genetics data</li>",
                              "<li>Communicate experimental results effectively, including writing research papers and giving presentations</li>",
                              # "<li>Effectively explain genetics concepts to different audiences</li>",
                              # "<li>Tap into the interdisciplinary nature of science</li>",
                              # "<li>Identify and critique scientific issues relating to society or ethics</li>",
                              "</ul>",
                              sep="\n"),
  
    "**Bioinformatics**",     paste(
                              "<ul>",
                              "<li>**C1**: Explain the role of computation and data mining in addressing hypothesis-driven and hypothesis-generating questions within the life sciences</li>",
                               # "<li>**C2**: Summarize key computational concepts, such as algorithms and relational databases, and their applications in the life sciences</li>",
                               # "<li>**C3**: Apply statistical concepts used in bioinformatics</li>",
                               "<li>**C4**: Use bioinformatics tools to examine complex biological problems in evolution, information flow, and other important areas of biology</li>",
                               "<li>**C5**: Find, retrieve, and organize various types of biological data</li>",
                               "<li>**C6**: Explore and/or model biological interactions, networks and data integration using bioinformatics</li>",
                               # "<li>**C7**: Use command-line bioinformatics tools and write simple computer scripts</li>",
                               # "<li>**C8**: Describe and manage biological data types, structure, and reproducibility</li>",
                               # "<li>**C9**: Interpret the ethical, legal, medical, and social implications of biological data</li>",
                              "</ul>",
                              sep="\n")
)

knitr::kable(competencies, col.names = NULL, format = "html", escape = FALSE)
```

Core concepts and competencies are taken from the following sources:

  - [Vision and Change in Undergraduate Biology Education](https://visionandchange.org/) AAAS report
  - [Genetics Learning Framework](https://genetics-gsa.org/education/genetics-learning-framework/) by [GSA](https://genetics-gsa.org/)
  - [Bioinformatics core competencies for undergraduate life sciences education](https://doi.org/10.1371/journal.pone.0196878) by [NIBLSE](https://qubeshub.org/community/groups/niblse)


### C-MOOR Collection {- .unlisted}

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "about_c-moor.Rmd",
  repo_name = "C-MOOR/C-MOOR_Template"
)
```
