---
title: "RNA-seq miniCURE"
date: "`r format(Sys.time(), '%B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Description about Course/Book."
favicon: assets/dasl_favicon.ico
output:
    bookdown::word_document2:
      toc: true
---

# About this Course {-}

This miniCURE allows students to develop a project using RNA-seq datasets to introduce how model organisms, high-throughput sequencing, and the scientific process are used in both basic and clinical research.

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1rWH7VTcPV1juH0E9NI-X6evMIKzgn1MQKlf_CRzT73w/edit#slide=id.g1f734f625ef_0_30")
```

### Skills Level {- .unlisted}

::: {.notice}
_Genetics_  
**Novice**: Introduction to [central dogma of molecular biology](https://openstax.org/books/biology-2e/pages/15-1-the-genetic-code)

_Programming skills_  
**Novice**: No programming experience needed
:::

### Learning Objectives {-}

- Learn about model organisms
- Explore high-throughput sequencing datasets
- Practice the [scientific process](https://undsci.berkeley.edu/understanding-science-101/how-science-works/)

### Core Competencies {-}

This activity addresses the following core concepts and competencies:

```{r, echo=FALSE}
competencies <- tibble::tribble(
  ~source, ~competencies,
  
  "**Vision and Change**",  paste(
                              "Concepts\n",
                              # "- **1**: EVOLUTION: The diversity of life evolved over time by processes of mutation, selection, and genetic change",
                              # "- **2**: STRUCTURE AND FUNCTION: Basic units of structure define the function of all living things",
                              "- **3**: INFORMATION FLOW, EXCHANGE, AND STORAGE: The growth and behavior of organisms are activated through the expression of genetic information in context",
                              # "- **4**: PATHWAYS AND TRANSFORMATIONS OF ENERGY AND MATTER: Biological systems grow and change by processes based upon chemical transformation pathways and are governed by the laws of thermodynamics",
                              # "- **5**: SYSTEMS: Living systems are interconnected and interacting.",
                              
                              "\nCompetencies\n",
                              "- **1**: Ability to apply the process of science",
                              "- **2**: Ability to use quantitative reasoning",
                              # "- **3**: Ability to use modeling and simulation",
                              # "- **4**: Ability to tap into the interdisciplinary nature of science",
                              "- **5**: Ability to communicate and collaborate with other disciplines",
                              # "- **6**: Ability to understand the relationship between science and society",
                            sep="\n"),

     "**Genetics**",         paste(
                              "\nConcepts\n",
                              # "- How is DNA organized?",
                              # "- What are the molecular components and mechanisms necessary to preserve and duplicate an organism’s genome?",
                              # "- What are the mechanisms by which an organism’s genome is passed on to the next generation?",
                              # "- How can one deduce information about genes, alleles, and gene functions from analysis of genetic crosses and patterns of inheritance?",
                              # "- How does the phenomenon of linkage affect the assortment of alleles during meiosis?",
                              "- How is genetic information expressed so it affects an organism’s structure and function?",
                              # "- How can gene activity be altered in the absence of DNA changes?",
                              # "- How do genes and genomes control changes in an organism’s structure and function throughout its life cycle?",
                              # "- How do different types of mutations affect genes and the corresponding mRNAs and proteins?",
                              # "- What are the processes that can affect the frequency of genotypes and phenotypes in a population over time?",
                              "- How do the results of molecular genetic studies in model organisms help us understand aspects of human genetics and genetic diseases?",
                              "- What experimental methods are commonly used to analyze gene structure, gene expression, gene function, and genetic variants?",
                              
                              "\nCompetencies\n",
                              "- Locate, read, and comprehend primary literature research papers on genetics topics",
                              "- Implement observational strategies to formulate a question",
                              "- Generate testable hypotheses",
                              # "- Design an experiment using appropriate controls and appropriate sample sizes",
                              # "- Gather and evaluate experimental evidence, including qualitative and quantitative data",
                              # "- Apply statistical methods when analyzing their data, and use patterns to construct a model",
                              "- Generate and interpret graphs displaying experimental results",
                              "- Critique large data sets and use bioinformatics to assess genetics data",
                              "- Communicate experimental results effectively, including writing research papers and giving presentations",
                              # "- Effectively explain genetics concepts to different audiences",
                              # "- Tap into the interdisciplinary nature of science",
                              # "- Identify and critique scientific issues relating to society or ethics",
                           sep="\n"),


  "**Bioinformatics**",      paste(
                               "- **C1**: Explain the role of computation and data mining in addressing hypothesis-driven and hypothesis-generating questions within the life sciences",
                               # "- **C2**: Summarize key computational concepts, such as algorithms and relational databases, and their applications in the life sciences",
                               # "- **C3**: Apply statistical concepts used in bioinformatics",
                               "- **C4**: Use bioinformatics tools to examine complex biological problems in evolution, information flow, and other important areas of biology",
                               "- **C5**: Find, retrieve, and organize various types of biological data",
                               "- **C6**: Explore and/or model biological interactions, networks and data integration using bioinformatics",
                               # "- **C7**: Use command-line bioinformatics tools and write simple computer scripts",
                               # "- **C8**: Describe and manage biological data types, structure, and reproducibility",
                               # "- **C9**: Interpret the ethical, legal, medical, and social implications of biological data",
                             sep="\n")

)

knitr::kable(competencies, col.names = NULL, format = "html")
```

Core concepts and competencies are taken from the following sources:

  - [Vision and Change in Undergraduate Biology Education](https://visionandchange.org/) AAAS report
  - [Genetics Learning Framework](https://genetics-gsa.org/education/genetics-learning-framework/) by [GSA](https://genetics-gsa.org/)
  - [Bioinformatics core competencies for undergraduate life sciences education](https://doi.org/10.1371/journal.pone.0196878) by [NIBLSE](https://qubeshub.org/community/groups/niblse)


### C-MOOR Collection {- .unlisted}

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "about_c-moor.Rmd",
  repo_name = "C-MOOR/C-MOOR_Template"
)
```
