---
title: "RNA-seq miniCURE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Description about Course/Book."
favicon: assets/dasl_favicon.ico
output:
    bookdown::gitbook:
      split_by: section
    bookdown::word_document2:
      toc: true
---

# About this Course {-}

This miniCURE allows students to develop a project using RNA-seq datasets to introduce how model organisms, high-throughput sequencing, and the scientific process are used in both basic and clinical research.

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1rWH7VTcPV1juH0E9NI-X6evMIKzgn1MQKlf_CRzT73w/edit#slide=id.g1f734f625ef_0_30")
```

### Skills Level {- .unlisted}

::: {.notice}
_Genetics_  
**Novice**: Introduction to [central dogma of molecular biology](https://openstax.org/books/biology-2e/pages/15-1-the-genetic-code)

_Programming skills_  
**Novice**: No programming experience needed
:::

### Learning Goals {-}

- **Navigate biological databases** to access the wealth of available scientific knowledge and data.
- **Analyze data (i.e. create and interpret plots)** to generate novel conclusions.
- **Cultivate an “-omics” perspective**, integrating multiple sources of information and wielding computational tools to enable a holistic approach
- **Practice the scientific process**, identifying avenues for research, designing experiments, analyzing data, and integrating results into the broader scientific discourse.

### Core Competencies {- .unlisted}

This activity addresses several core concepts and competencies from the following sources:

  - [Vision and Change in Undergraduate Biology Education](https://visionandchange.org/) AAAS report
  - [Genetics Learning Framework](https://genetics-gsa.org/education/genetics-learning-framework/) by [GSA](https://genetics-gsa.org/)
  - [Bioinformatics core competencies for undergraduate life sciences education](https://doi.org/10.1371/journal.pone.0196878) by [NIBLSE](https://qubeshub.org/community/groups/niblse)

See [Appendix](#competencies-table) for details.


### C-MOOR Collection {- .unlisted}

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "about_c-moor.Rmd",
  repo_name = "C-MOOR/miniCURE-RNA-seq"
)
```

