# (PART\*) Exploration and Discovery {-}

# Scientific Literature

## Pre-lab {#pre-lab-scientific-literature}

Familiarize yourself with the paper "[A high-sugar diet produces obesity and insulin resistance in wild-type Drosophila](https://pubmed.gov/21719444)" by  Musselman et al., 2011 Dis Model Mech

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1rWH7VTcPV1juH0E9NI-X6evMIKzgn1MQKlf_CRzT73w/edit#slide=id.g1f83ffdfb86_0_52")
```

## Lab Slides {#lab-slides-scientific-literature}

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1uuuAbg_rcfCVohaVrrxfUfszwTDqigzg34qEWfTNeR0/edit#slide=id.g35f391192_00")
```

[[slides](https://docs.google.com/presentation/d/1uuuAbg_rcfCVohaVrrxfUfszwTDqigzg34qEWfTNeR0)]

## Lab Activity: Scientific Literature

### Instructions

Follow your instructor’s instructions  and answer the following questions with your group. In your responses, properly paraphrase (do not copy/paste or inappropriately paraphrase).  To receive credit, you must use your own words and sentence structure.  Using the same sentence structure (and changing a few words only) is considered plagiarism.

[Link to Musselman et al. 2011](https://journals.biologists.com/dmm/article/4/6/842/3157/A-high-sugar-diet-produces-obesity-and-insulin). You can also download a pdf from the website.

### Part 1. Overview of the paper

Determine the main objectives and purpose of the paper.  Re-read the Abstract (or Summary) and the introduction with your group.

| 1. What is the purpose of the study? |
|:-|
| <br> |

<br>

| 2. What is the hypothesis? |
|:-|
| <br> |

<br>

| 3. Describe the knowledge gap. In essence, what did the scientific community not know that this study was trying to answer? |
|:-|
| <br> |

### Part 2. Figure Analysis - Methods and Findings

**Methods.** How did the researchers test their hypothesis? Explain in your own words the methods in each figure.

| Figure | Methods |
|:-|:-|
| Fig. 1A-F | |
| Fig. 2A-D | |
| Fig. 2E-2I | |
| Fig. 3A-E | |
| Fig. 4A-B | |

**Results.** What are the main findings from each figure?

| Figure | Main Findings |
|:-|:-|
| Fig. 1A-F | |
| Fig. 2A-D | |
| Fig. 2E-2I | |
| Fig. 3A-E | |
| Fig. 4A-B | |

### Part 3.  Your Conclusions

| 1. Work with your group and write 3-4 sentences summarizing the main findings/conclusions of the paper based one what you came up with in Part 3. |
|:-|
| <br> |

<br>

| 2. Talk with your group and come up with a list of 2-3 limitations. |
|:-|
| <br> |

### Part 4. The Author’s Conclusions (Discussion Section)

| 1. Read the discussion section and compare it to what your group came up with in Part 4. How similar are the conclusions you came up with to the authors? Were there any differences in interpreting the data? Different conclusions based on those data? |
|:-|
| <br> |

<br>

| 2. What do you think about the conclusions made by the authors? Do you think they are over-stating their findings? Do you think their conclusions are accurate and appropriate? |
|:-|
| <br> |

### Part 5. The Future

| 1. Scientific work builds on previous studies. What do you believe could be the next step to further the work these researchers did? |
|:-|
| <br> |

<br>

| 2. Do you believe further research in this area may benefit society? Can we build on what this study found? |
|:-|
| <br> |

### Footnotes

#### Resources

- [Google Doc](https://docs.google.com/document/d/1QmuNMHKiHRGXohS0P9Pojc3zBvUZDG8L)

#### Contributions and Affiliations

- Stephanie Coffman, Clovis Community College

Last Revised: February 15, 2022
